### 2024-09-24 paper/paper.qmd
Initial commit

---
title: "Investigating Bicycle Theft in Toronto: Trends and Patterns"
subtitle: "."
author: 
  - Taejun Um
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "In urban environments like Toronto, bicycle theft is as common as in any major city worldwide. This paper examines the trends and patterns of bicycle theft in Toronto from 2014 to 2022, focusing on the cost of bicycles and the types of locations where the thefts occurred. Data from this period show that the overall number of bicycle thefts did not significantly decline but began to decrease gradually from 2020, particularly in thefts occurring at residential properties and those involving high-end bicycles. These findings can inform community efforts and assist Toronto's public authorities in developing more effective strategies to protect citizens' property and reduce bicycle theft."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(opendatatoronto)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

Toronto is the most populous and densely populated city in Canada. As a result, traffic congestion makes it difficult for people to navigate the city, and many citizens report this as a significant issue*. This has led to an increase in the number of cyclists using bicycles for efficient point-to-point travel within the city. Currently, approximately 659 km of cycling routes are interwoven like a spider's web throughout Toronto*. Unfortunately, bicycle thefts remain a persistent problem for residents.

This paper examines reported bicycle thefts in Toronto from 2014 to 2022. The main objective is to observe and analyze the number of thefts, the locations of the incidents, and the price range of the stolen bicycles. Specifically, we will compare trends in the locations of incidents by year alongside changes in the price range of the targeted bicycles. Through this investigation, I confirmed that thefts steadily increased from 2014 until 2020, when the number of thefts began to decline. Additionally, there have been notable changes in the locations where thefts occurred and the price range of the stolen bicycles over the years. Understanding these trends and patterns will help prevent future bicycle thefts, and inform new measures by Toronto’s public authorities to enhance theft prevention.

The remainder of this paper is structured as follows: @sec-data discusses the data and process of data cleaning, and @sec-first-point discusses our findings, limitation of the study, and suggestions to reduce bicycle theft.

# Data {#sec-data}

Some of our data is of penguins (), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

And also planes (). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false
date_counts <- as.data.frame(table(data$occ_date))
colnames(date_counts) <- c("date", "count")
ggplot(date_counts, aes(x=as.Date(date), y=count)) + geom_line()

```

Talk way more about it.

```{r}
data |>
  ggplot(mapping = aes(x=occ_year, fill = premises_type)) + geom_bar(position = "dodge2")

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

\newpage

# References

### 2024-09-24 scripts/00-simulate_data.R

@@ -1,0 +1,51 @@
+#### Preamble ####
+# Purpose: Simulates... [...UPDATE THIS...]
+# Author: Taejun Um
+# Date: 24 September 2024
+# Contact: taejun.um@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Simulate data ####
+set.seed(4020)
+
+number_of_occurrences <- 100
+
+# Define the range of dates
+start_date <- as.Date("2014-01-01")  # Start date
+end_date <- as.Date("2022-12-31")    # End date
+
+# Create a sequence of dates between the start and end date
+date_sequence <- seq(from = start_date, to = end_date, by = "day")
+sample_date <- sample(date_sequence, 100)
+year_value <- format(sample_date, "%Y")
+
+#Create a set of premises
+sample_premises <- c("Apartment", "Commercial", "Educational", "House", "Other",
+                     "Outside", "Transit")
+
+simulated_data <-
+  tibble(
+    id = 1:100,
+    occ_date = sample_date,
+    occ_year = year_value,
+    premises_type = sample(sample_premises, 100, TRUE),
+    bike_cost = rgamma(100, 3, 0.005),
+    cost_group =
+      case_when(
+        bike_cost < 500 ~ "<500",
+        bike_cost < 1000 ~ "500<1000",
+        bike_cost < 1500 ~ "1000<1500",
+        bike_cost < 2000 ~ "1500<2000",
+        bike_cost >= 2000 ~ "2000<"
+      )
+    )
+
+####Write CSV
+write_csv(simulated_data, file = "data/raw_data/simulated.csv")

### 2024-09-24 scripts/01-download_data.R

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from Open Data Toronto.
+# Author: Taejun Um
+# Date: 24 September 2024
+# Contact: taejun.um@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(dplyr)
+library(tidyverse)
+
+
+#### Download data ####
+
+
+# get package
+package <- show_package("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
+package
+
+# get all resources for this package
+resources <- list_package_resources("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+the_raw_data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+the_raw_data
+
+#### Save data ####
+write_csv(the_raw_data, "data/raw_data/raw_data.csv") 
+
+         

### 2024-09-24 scripts/02-data_cleaning.R

@@ -1,0 +1,44 @@
+#### Preamble ####
+# Purpose: Cleans the raw data and add new variable to categorize bike cost.
+# Author: Taejun Um
+# Date: 24 September 2024
+# Contact: taejun.um@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None.
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+
+#### Clean data ####
+raw_data <- read_csv("data/raw_data/raw_data.csv")
+
+cleaned_data <-
+  raw_data|>
+  select("_id", OCC_DATE, OCC_YEAR, PREMISES_TYPE, BIKE_COST)|>
+  janitor::clean_names()
+
+cleaned_data <-
+  cleaned_data |>
+  mutate(
+    cost_group =
+      case_when(
+        bike_cost < 500 ~ "<500",
+        bike_cost < 1000 ~ "500<1000",
+        bike_cost < 1500 ~ "1000<1500",
+        bike_cost < 2000 ~ "1500<2000",
+        bike_cost >= 2000 ~ "2000<"
+      ),
+    cost_group = 
+      factor(cost_group, levels = c("<500", "500<1000", "1000<1500", "1500<2000", "2000<"))
+  )
+
+cleaned_data<-
+  cleaned_data|>
+  filter(occ_year > 2013, occ_year < 2023)|> 
+  tidyr::drop_na()
+
+
+#### Save data ####
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-24 scripts/03-test_data.R

@@ -1,0 +1,33 @@
+#### Preamble ####
+# Purpose: Tests... [...UPDATE THIS...]
+# Author: Taejun Um
+# Date: 24 September 2024
+# Contact: taejun.um@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have simulated data
+# Any other information needed? None.s
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Test data ####
+data <- read_csv("data/raw_data/simulated.csv")
+
+#Test1 Check occ_date is in date form
+"Date" == class(data$occ_date)
+
+#Test2 Check occ_year has a proper year
+all(format(data$occ_date, "%Y") == data$occ_year)
+
+#Test3 Check if premises_types have correct value
+premises_test <- c("Apartment", "Commercial", "Educational", "House", "Other",
+                   "Outside", "Transit")
+all(data$premises_type %in% premises_test)
+
+#Test4 Check if bike_cost is correctly assigned for each observation
+data$bike_cost
+
+#Test5 Test for NAs
+all(!is.na(data))

### 2024-09-24 paper/paper.qmd
Edited Discussion and graphs (first draft)

@@
 library(opendatatoronto)
+library(knitr)
 
@@
 # Data {#sec-data}
+## Overview {#sec-overview}
 
-Some of our data is of penguins (), from @palmerpenguins.
+The dataset used in the paper covers bicycle thefts reported to the Toronto Police system between 2014 and 2024.
+This dataset has been updated as of June 2024 and is considered reliable as it is based on the Toronto Police system.
+
+The dataset was downloaded from OpenDataToronto, and its setup was completed with the help of tidyverse, janitor, and tidyr.
+
+The original dataset contained 28 variables, but I used the following: ID, occurrence date, occurrence year, premises type, and bike cost.
+Then, I created a new group called "cost group" based on the bike cost, categorizing the bicycles according to their price range.
+Since the dataset is based on the date the theft was reported, I organized the data by the actual date of occurrence to reflect events from 2014 to 2022, excluding incidents that occurred before 2014.
+Additionally, during the data cleaning phase, observations with missing information were excluded.
+As a result, the dataset used in this paper contains 28,293 observations.
+(@tbl-head) shows an first 5 row of the data used.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-head
+#| tbl-cap: First 5 rows of data
+#| warning: false
+#| message: false
 #| echo: false
+
+data |>
+  slice(1:5) |>
+  kable(
+    digits = 4,
+    booktabs = TRUE,
+    linesep = ""
+  )
+```
+
+## Number of Occurrence
+
+First, we will examine how many bicycle thefts occurred by date.
+@fig-occur is a line chart that records thefts on a daily basis.
+Bicycle thefts showed an increasing trend from 2014, followed by a noticeable decrease starting in 2020.
+
+There were also clear differences by season.
+Each year, there was a sharp increase in thefts during the middle of the year, while the beginning and end of the year saw a sharp decline, showing a cyclical pattern.
+This indicates that thefts significantly increased during the summer season, while they were almost non-existent in the winter compared to summer.
+
+```{r}
+#| label: fig-occur
+#| fig-cap: Number of bicycle theft incidents
+#| warning: false
+#| message: false
+#| echo: false
+
+date_counts <- as.data.frame(table(data$occ_date))
+colnames(date_counts) <- c("date", "count")
+date_counts|>
+  ggplot(aes(x = as.Date(date), y = count)) +
+  geom_line() +
+  theme_minimal() +
+  labs(x = "Date", y = "Number of Occurrences")
 
@@
 
-Talk more about it.
+## Types of Crime Scenes
 
-And also planes ().
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+In this dataset, crime scenes have been categorized into seven types: Apartment, Commercial, Educational, House, Outside, Transit, and Others.
+As seen in @fig-crime-scenes, there has been significant fluctuation in crimes committed in residential areas.
+Residential areas are classified into two categories: apartments and houses.
+Among them, apartment-related thefts showed a steady linear increase until 2019, but in 2020, the number dropped by around 200 incidents in one year, returning to 2017 levels by 2022.
+House-related thefts followed a zigzag pattern, increasing from 2014 and reaching a peak in 2020, before dropping to less than half by 2022.
+
+Another interesting point is that thefts committed outdoors were more frequent than those in apartments before 2019.
+However, starting in 2019, apartment thefts surpassed outdoor thefts, until 2022, when outdoor thefts once again became more common than apartment thefts.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: fig-crime-scenes
+#| fig-cap: Number of occurrences categorized by crime scene
 #| echo: false
@@
 #| message: false
-date_counts <- as.data.frame(table(data$occ_date))
-colnames(date_counts) <- c("date", "count")
-ggplot(date_counts, aes(x=as.Date(date), y=count)) + geom_line()
+data |>
+  ggplot(mapping = aes(x=occ_year, fill = premises_type)) + 
+  geom_bar(position = "dodge2") +
+  labs(x = "Year", y = "Number of Occurrences", fill = "Types of Crime Scenes") +
+  theme(legend.position = "bottom")
 
@@
 
-Talk way more about it.
+## Cost of Stolen Bicycle
 
+The original dataset includes a variable for the price of stolen bicycles, but it does not contain a variable for grouped prices.
+Therefore, I created a new variable called cost_group by grouping the prices in increments of $500.
+@fig-stolen-bike shows a bar graph with number of bike stolen by year categorized by cost group.
+From 2020 to 2022, the thefts of relatively inexpensive bicycles, priced under $1,000, have significantly decreased.
+On the other hand, high-end bicycles priced over $2,000 were stolen about 250 times in 2014.
+But this number has steadily increased, reaching approximately 500 thefts by 2022.
 ```{r}
+#| label: fig-stolen-bike
+#| fig-cap: Total cost of bicycle by year
+#| fig-pos: "H"
+#| echo: false
+#| warning: false
+#| message: false
 data |>
-  ggplot(mapping = aes(x=occ_year, fill = premises_type)) + geom_bar(position = "dodge2")
+  ggplot(mapping = aes(x=occ_year, fill = cost_group)) + 
+  geom_bar(position = "dodge2") +
+  labs(x = "Year", y = "Cost of Bicycle", fill = "Cost Group") +
+  theme(legend.position = "bottom")
 
@@
 
-## First discussion point {#sec-first-point}
+## Overall {#sec-first-point}
+The analysis of bicycle theft in Toronto has revealed trends and patterns over the nine years of theft incidents.
+All trends shifted starting in 2020, which appears to be a result of the lockdowns due to the pandemic.
+First, the overall number of thefts began to decrease from 2020, indicating a significant reduction in outdoor activities.
+Bicycle owners were riding their bikes less frequently, and thieves were also unable to engage in outdoor activities due to the lockdowns.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+One of the most notable trends is that the daily theft graph exhibits a cyclical pattern.
+This indicates that most crimes occur in the summer, while very few happen in the winter.
+The winter in Toronto, with its snow and cold weather, makes cycling difficult for both owners and thieves, which likely explains the lack of activity in both groups during this season.
 
-## Second discussion point
+When examining thefts by location, we observed significant changes in residential area thefts.
+With the increase in remote work and online learning, more people have been at home, leading to a substantial decrease in thefts within residential spaces.
 
-## Third discussion point
+There were also intriguing observations regarding the price range of stolen bicycles.
+Year after year, the number of thefts of more expensive bicycles has been on the rise.
+This can be interpreted in two ways: either bicycle thieves have started to specifically target high-end bicycles, or the prices of bicycles available on the market have increased.
+What is clear is that the rising theft rates of expensive bicycles have led to more severe property losses for citizens.
+Bicycle theft may be considered a minor issue, but it is a clear crime that deprives someone of their valuable property or an essential means of transportation for their life.
+
+## Implication
+
+According to this data, the Toronto Police should strengthen patrols and response efforts during the summer to prevent bicycle thefts.
+As of the data up to 2022, thefts in residential areas have significantly decreased; however, with the end of the lockdown, there is a possibility that these incidents may rise again.
+Therefore, it would be beneficial to conduct preventive campaigns targeting residential areas.
+
+Regardless of their price, all bicycles are important, but there has been a recent trend of increasing thefts of high-end bicycles.
+To protect citizens' property, it is essential to deliver messages of caution to the community.
 
@@
 
-Weaknesses and next steps should also be included.
+This dataset has a significant amount of missing data, and unfortunately, I had to drop some parts of the dataset.
+Additionally, only theft incidents that were reported are included, meaning that incidents that were not reported for various reasons are not accounted for.
+These missing pieces of information may have influenced the trends or patterns identified in this paper.
 

### 2024-09-25 scripts/01-download_data.R
fixing download_date to get full data from opendatatoronto

@@ -18,20 +18,19 @@ library(tidyverse)
 
 
 # get package
-package <- show_package("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
-package
-
-# get all resources for this package
-resources <- list_package_resources("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-the_raw_data <- filter(datastore_resources, row_number()==1) %>% get_resource()
-the_raw_data
-
-#### Save data ####
-write_csv(the_raw_data, "data/raw_data/raw_data.csv") 
-
+# package <- show_package("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
+# package
+# 
+# # get all resources for this package
+# resources <- list_package_resources("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
+# 
+# # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+# datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+# 
+# # load the first datastore resource as a sample
+# the_raw_data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+# the_raw_data
+# 
+# #### Save data ####
+# write_csv(the_raw_data, "data/raw_data/raw_data.csv") 
          

### 2024-09-26 paper/paper.qmd
Edited cleaned dataset and some graphs in paper

@@
 #| echo: false
+#| layout-ncol: 2
 
-date_counts <- as.data.frame(table(data$occ_date))
-colnames(date_counts) <- c("date", "count")
-date_counts|>
-  ggplot(aes(x = as.Date(date), y = count)) +
+date_counts <- data |>
+  group_by(occ_year) |>
+  summarise(total_occurrences = n())
+
+ggplot(date_counts, mapping = aes(x = occ_year, y = total_occurrences)) +
   geom_line() +
@@
   labs(x = "Date", y = "Number of Occurrences")
+
+date_counts |>
+  slice(1:10) |>
+  kable(
+    digits = 4,
+    booktabs = TRUE,
+    linesep = ""
+  )
 
@@
   geom_bar(position = "dodge2") +
+  facet_wrap(~ premises_type) +
   labs(x = "Year", y = "Number of Occurrences", fill = "Types of Crime Scenes") +
-  theme(legend.position = "bottom")
+  theme(
+    legend.text = element_text(size = 7),
+    legend.title = element_text(size = 9),
+    legend.key.size = unit(0.4, "cm"),legend.position = "bottom"
+    )
 
@@
 #| message: false
+
+data <-
+  data|>mutate(
+  cost_group =
+      factor(cost_group, levels = c("< $500", "$500 < $1000", "$1000 < $1500", "$1500 < $2000", "$2000 <")))
 data |>
   ggplot(mapping = aes(x=occ_year, fill = cost_group)) + 
-  geom_bar(position = "dodge2") +
+  geom_bar() +
+  facet_wrap(~ cost_group) +
   labs(x = "Year", y = "Cost of Bicycle", fill = "Cost Group") +
-  theme(legend.position = "bottom")
+  theme(
+    legend.text = element_text(size = 7),
+    legend.title = element_text(size = 9),
+    legend.key.size = unit(0.4, "cm"),
+    legend.position = "bottom"
+  )
+```
+
+```{r}
 

### 2024-09-26 scripts/01-download_data.R

@@ -13,6 +13,8 @@ library(opendatatoronto)
 library(dplyr)
 library(tidyverse)
 
+#Following codes are direct copy from opendata website. However, it only downloads first 32000 rows of observations.
+#Therefore, I decided to use read_csv function save full datasets.
 
 #### Download data ####
 
@@ -33,4 +35,7 @@ library(tidyverse)
 # 
 # #### Save data ####
 # write_csv(the_raw_data, "data/raw_data/raw_data.csv") 
-         
+
+#Downloading all data from opendatatoronto directly to get all observations.
+all_data <- read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/c7d34d9b-23d2-44fe-8b3b-cd82c8b38978/resource/55136dac-26b1-4028-b9f5-7c2344f94153/download/bicycle-thefts%20-%204326.csv")
+write_csv(all_data, "data/raw_data/raw_data.csv")

### 2024-09-26 scripts/02-data_cleaning.R

@@ -24,20 +24,33 @@ cleaned_data <-
   mutate(
     cost_group =
       case_when(
-        bike_cost < 500 ~ "<500",
-        bike_cost < 1000 ~ "500<1000",
-        bike_cost < 1500 ~ "1000<1500",
-        bike_cost < 2000 ~ "1500<2000",
-        bike_cost >= 2000 ~ "2000<"
+        bike_cost < 500 ~ "< $500",
+        bike_cost < 1000 ~ "$500 < $1000",
+        bike_cost < 1500 ~ "$1000 < $1500",
+        bike_cost < 2000 ~ "$1500 < $2000",
+        bike_cost >= 2000 ~ "$2000 <"
       ),
     cost_group = 
-      factor(cost_group, levels = c("<500", "500<1000", "1000<1500", "1500<2000", "2000<"))
-  )
+      factor(cost_group, levels = c("< $500", "$500 < $1000", "$1000 < $1500", "$1500 < $2000", "$2000 <"))
+  ) 
+
 
 cleaned_data<-
   cleaned_data|>
-  filter(occ_year > 2013, occ_year < 2023)|> 
+  filter(occ_year > 2013, occ_year < 2024)|> 
   tidyr::drop_na()
+
+cleaned_data <-
+  cleaned_data|>mutate(
+  cost_group =
+      factor(cost_group, levels = c("< $500", "$500 < $1000", "$1000 < $1500", "$1500 < $2000", "$2000 <")))
+
+cleaned_data |>
+  ggplot(mapping = aes(x=occ_year, fill = cost_group)) + 
+  geom_bar(position = "dodge2") +
+  labs(x = "Year", y = "Cost of Bicycle", fill = "Cost Group") +
+  theme(legend.position = "bottom")
+
 
 
 #### Save data ####

### 2024-09-26 paper/paper.qmd
Improved appearances of graphs

@@
 
-The original dataset contained 28 variables, but I used the following: ID, occurrence date, occurrence year, premises type, and bike cost.
+The original dataset contained 28 variables, but I used the following: ID, occurrence date, occurrence year, occurrence month, premises type, and bike cost.
 Then, I created a new group called "cost group" based on the bike cost, categorizing the bicycles according to their price range.
@@
 Additionally, during the data cleaning phase, observations with missing information were excluded.
-As a result, the dataset used in this paper contains 28,293 observations.
+As a result, the dataset used in this paper contains 30,937 observations.
 (@tbl-head) shows an first 5 row of the data used.
@@
 
-## Number of Occurrence
+## Number of Occurrence by Year
 
@@
   geom_line() +
+  geom_point(size = 2) +
   theme_minimal() +
-  labs(x = "Date", y = "Number of Occurrences")
+  labs(x = "Year", y = "Number of Occurrences")
 
@@
   kable(
+    col.names = c("Year", "Number of Occurrences"),
     digits = 4,
@@
 
+```
+
+## Seasonal Effect
+
+```{r}
+#| label: fig-month
+#| fig-cap: Number of bicycle theft incidents by month
+#| warning: false
+#| message: false
+#| echo: false
+
+data$occ_month <- 
+  factor(data$occ_month, levels = c("January", "February", "March", "April", "May", "June", 
+                                        "July", "August", "September", "October", "November", "December"))
+
+month_grouped <- data |>
+  group_by(occ_year, occ_month) |>
+  summarise(number_of_observations = n(), .groups = "drop")
+
+ggplot(month_grouped, mapping = aes(x = occ_month, y = number_of_observations, color = factor(occ_year), group = occ_year)) +
+  geom_line(size = 0.5) +
+  geom_point(size = 2) +
+  scale_color_brewer(palette = "Set1") +
+  labs(x = "Month", y = "Number of Occurrences", color = "Year") +
+  theme_minimal() +
+  theme(text = element_text(size = 12), axis.text.x = element_text(angle = 90)) +
+  scale_color_brewer(palette = "Set3")
+
+```
+
+## Time Effect
+
+```{r}
+#| label: fig-hour
+#| fig-cap: Number of bicycle theft incidents by hour in 10 years
+#| warning: false
+#| message: false
+#| echo: false
+
+time_grouped <- data |>
+  group_by(occ_hour) |>
+  summarise(number_of_observations = n())
+
+ggplot(time_grouped, mapping = aes(x = occ_hour, y = number_of_observations)) +
+  geom_line(size = 0.5) +
+  geom_point(size = 2) +
+  labs(x = "Hour", y = "Number of Occurrences") +
+  theme_minimal() +
+  theme(text = element_text(size = 12), axis.text.x = element_text(angle = 90))
 ```
@@
 
-```{r}
-
-```
-
 # Discussion

### 2024-09-26 scripts/02-data_cleaning.R

@@ -16,7 +16,7 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 
 cleaned_data <-
   raw_data|>
-  select("_id", OCC_DATE, OCC_YEAR, PREMISES_TYPE, BIKE_COST)|>
+  select("_id", OCC_DATE, OCC_YEAR, OCC_MONTH, OCC_HOUR, PREMISES_TYPE, BIKE_COST)|>
   janitor::clean_names()
 
 cleaned_data <-
@@ -39,19 +39,15 @@ cleaned_data<-
   cleaned_data|>
   filter(occ_year > 2013, occ_year < 2024)|> 
   tidyr::drop_na()
-
-cleaned_data <-
-  cleaned_data|>mutate(
-  cost_group =
-      factor(cost_group, levels = c("< $500", "$500 < $1000", "$1000 < $1500", "$1500 < $2000", "$2000 <")))
-
-cleaned_data |>
-  ggplot(mapping = aes(x=occ_year, fill = cost_group)) + 
-  geom_bar(position = "dodge2") +
-  labs(x = "Year", y = "Cost of Bicycle", fill = "Cost Group") +
-  theme(legend.position = "bottom")
-
-
+# 
+# cleaned_data <-
+#   cleaned_data|>mutate(
+#   cost_group =
+#       factor(cost_group, levels = c("< $500", "$500 < $1000", "$1000 < $1500", "$1500 < $2000", "$2000 <")))
+# 
+# cleaned_data$occ_month <- 
+#   factor(cleaned_data$occ_month, levels = c("January", "February", "March", "April", "May", "June", 
+#                                         "July", "August", "September", "October", "November", "December"))
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-26 paper/paper.qmd
updated links to repo and citations

@@
 title: "Investigating Bicycle Theft in Toronto: Trends and Patterns"
-subtitle: "."
 author: 
   - Taejun Um
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/tonytto/Investigating-Bicycle-Theft-in-Toronto"
 date: today
@@
 library(tidyverse)
-library(ggplot2)
 library(opendatatoronto)
@@
 Toronto is the most populous and densely populated city in Canada.
-As a result, traffic congestion makes it difficult for people to navigate the city, and many citizens report this as a significant issue*.
+As a result, traffic congestion makes it difficult for people to navigate the city, and many citizens report this as a significant issue [@traffic].
 This has led to an increase in the number of cyclists using bicycles for efficient point-to-point travel within the city.
-Currently, approximately 659 km of cycling routes are interwoven like a spider's web throughout Toronto*.
+Currently, approximately 659 km of cycling routes are interwoven like a spider's web throughout Toronto [@bikelane].
 Unfortunately, bicycle thefts remain a persistent problem for residents.
@@
 
-The dataset was downloaded from OpenDataToronto, and its setup was completed with the help of tidyverse, janitor, and tidyr.
+The dataset was downloaded from Toronto's Open Data Portal using opendatatoronto package [@opendatatoronto].
+Its setup was completed through R [@citeR] with the help of tidyverse package [@citetidyverse], janitor package [@citejanitor] and knitr package [@citeknitr].
 
-The original dataset contained 28 variables, but I used the following: ID, occurrence date, occurrence year, occurrence month, premises type, and bike cost.
+The original dataset contained 28 variables, but I used the following: ID, occurrence date, occurrence year, occurrence month, occurrence hour, premises type, and bike cost.
 Then, I created a new group called "cost group" based on the bike cost, categorizing the bicycles according to their price range.
@@
 Additionally, during the data cleaning phase, observations with missing information were excluded.
-As a result, the dataset used in this paper contains 30,937 observations.
+As a result, the dataset used in this paper contains 29,217 observations.
 (@tbl-head) shows an first 5 row of the data used.

### 2024-09-26 scripts/01-download_data.R
erased double loaded packages

@@ -10,7 +10,6 @@
 
 #### Workspace setup ####
 library(opendatatoronto)
-library(dplyr)
 library(tidyverse)
 
 #Following codes are direct copy from opendata website. However, it only downloads first 32000 rows of observations.

### 2024-09-26 scripts/02-data_cleaning.R

@@ -9,7 +9,6 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(dplyr)
 
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/raw_data.csv")

### 2024-09-26 scripts/02-data_cleaning.R
Edited data cleaning process

@@ -38,15 +38,6 @@ cleaned_data<-
   cleaned_data|>
   filter(occ_year > 2013, occ_year < 2024)|> 
   tidyr::drop_na()
-# 
-# cleaned_data <-
-#   cleaned_data|>mutate(
-#   cost_group =
-#       factor(cost_group, levels = c("< $500", "$500 < $1000", "$1000 < $1500", "$1500 < $2000", "$2000 <")))
-# 
-# cleaned_data$occ_month <- 
-#   factor(cleaned_data$occ_month, levels = c("January", "February", "March", "April", "May", "June", 
-#                                         "July", "August", "September", "October", "November", "December"))
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-26 paper/paper.qmd
Edited first part of data analysis

@@
 # Data {#sec-data}
-## Overview {#sec-overview}
+## Overview {#sec-data-overview}
 
@@
 
-## Number of Occurrence by Year
+## Number of Occurrences by Time
 
-First, we will examine how many bicycle thefts occurred by date.
-@fig-occur is a line chart that records thefts on a daily basis.
-Bicycle thefts showed an increasing trend from 2014, followed by a noticeable decrease starting in 2020.
-
-There were also clear differences by season.
-Each year, there was a sharp increase in thefts during the middle of the year, while the beginning and end of the year saw a sharp decline, showing a cyclical pattern.
-This indicates that thefts significantly increased during the summer season, while they were almost non-existent in the winter compared to summer.
+First, we will examine how many bicycle thefts occurred by year.
+In @fig-occur-year-1, the line chart records thefts yearly and the table records the number of incidents in numbers.
+Bicycle thefts showed an increasing trend from 2014, peaking in 2018 with 3,418 recorded incidents, followed by a noticeable decrease starting in 2021.
+The decrements that happened in 2021 continued through 2023.
+According to @fig-occur-year-2, the number of incidents recorded in 2022 was 2,491 which was fewer than 2,511 incidents recorded in 2014.
+This indicates that the trend of bicycle theft has been decreasing over the last three years.
 
 ```{r}
-#| label: fig-occur
-#| fig-cap: Number of bicycle theft incidents
+#| label: fig-occur-year
+#| fig-cap: Number of bicycle theft incidents by year
 #| warning: false
@@
 #| echo: false
-#| layout-ncol: 2
 
@@
 
+```
+```{r}
+#| label: tbl-occur-year
+#| tbl-cap: Number of bicycle theft incidents by year
+#| warning: false
+#| message: false
+#| echo: false
+
 date_counts |>
@@
   )
-
 ```
 
-## Seasonal Effect
+@fig-month shows the number of incidents that occurred by month with each line representing a different year.
+This graph indicates there are clear differences by season.
+Each year, there was a sharp increase in thefts during the middle of the year, while the beginning and end of the year saw a sharp decline, showing a cyclical pattern.
+The number of incidents peaked in the summer months (June to August) and the incidents in January and December dropped to less than one-third compared to the summer months in all years.
 
@@
 #| echo: false
+#| fig-pos: "H"
 
@@
 
-## Time Effect
+### By hour
+
+@fig-hour represents the number of bicycle theft incidents over an hour.
+@fig-hour-1 shows the average number of incidents across all years, while the @fig-hour-2 illustrates the number of incidents for each year over an hour.
+Graph 1 shows that there is a steady increase in thefts from 5 AM to 6 PM, and after that, the numbers begin to decline.
+Notably, there are three distinct peaks at 9 AM, 12 PM, and 6 PM, corresponding to periods of high social activity and movement.
+On average over the past decade, the lowest number of incidents occurred at 5 AM, with 37.3 thefts reported annually, whereas 6 PM saw the highest, with 218 thefts per year.
+From 1 AM to 7 AM, fewer than 100 incidents were recorded each year, but beyond that, the number consistently exceeded 100.
 
@@
 #| fig-cap: Number of bicycle theft incidents by hour in 10 years
+#| fig-subcap: ["Average in Year","Separated in Each Year"]
 #| warning: false
@@
 #| echo: false
+#| layout-ncol: 2
+#| fig-pos: "H"
 
@@
   group_by(occ_hour) |>
-  summarise(number_of_observations = n())
+  summarise(number_of_observations = n()/10)
 
@@
   labs(x = "Hour", y = "Number of Occurrences") +
+  theme_minimal()
+
+time_grouped_sep <- data |>
+  group_by(occ_year, occ_hour) |>
+  summarise(number_of_observations_h = n(), .groups = "drop")
+
+ggplot(time_grouped_sep, mapping = aes(x = occ_hour, y = number_of_observations_h, color = factor(occ_year), group = occ_year)) +
+  geom_line(size = 0.5) +
+  geom_point(size = 2) +
+  scale_color_brewer(palette = "Set1") +
+  labs(x = "Month", y = "Number of Occurrences", color = "Year") +
   theme_minimal() +
-  theme(text = element_text(size = 12), axis.text.x = element_text(angle = 90))
+  scale_color_brewer(palette = "Set3")
 ```
@@
 #| message: false
+#| fig-pos: "H"
+
 data |>

### 2024-09-27 scripts/00-simulate_data.R
Added new variables and fixed preamble

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
+# Purpose: Simulates Bicycle Theft data for Toronto from 2014 to 2023
 # Author: Taejun Um
-# Date: 24 September 2024
+# Date: 27 September 2024
 # Contact: taejun.um@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
@@ -10,7 +10,6 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Simulate data ####
 set.seed(4020)
@@ -29,14 +28,19 @@ year_value <- format(sample_date, "%Y")
 #Create a set of premises
 sample_premises <- c("Apartment", "Commercial", "Educational", "House", "Other",
                      "Outside", "Transit")
+sample_months <- sample(c("January", "February", "March", "April", "May", "June", 
+                   "July", "August", "September", "October", "November", "December"), 100, replace = TRUE)
+sample_hour <- sample(1:23, 100, replace = TRUE)
 
 simulated_data <-
   tibble(
     id = 1:100,
     occ_date = sample_date,
     occ_year = year_value,
+    occ_month = sample_months,
+    occ_hour = sample_hour,
     premises_type = sample(sample_premises, 100, TRUE),
-    bike_cost = rgamma(100, 3, 0.005),
+    bike_cost = round(rgamma(100, 3, 0.005), 2),
     cost_group =
       case_when(
         bike_cost < 500 ~ "<500",

### 2024-09-27 scripts/01-download_data.R
Erased unused line

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from Open Data Toronto.
 # Author: Taejun Um
-# Date: 24 September 2024
+# Date: 27 September 2024
 # Contact: taejun.um@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
@@ -12,28 +12,10 @@
 library(opendatatoronto)
 library(tidyverse)
 
-#Following codes are direct copy from opendata website. However, it only downloads first 32000 rows of observations.
+#Codes from opendata website only downloads first 32000 rows of observations.
 #Therefore, I decided to use read_csv function save full datasets.
 
 #### Download data ####
-
-
-# get package
-# package <- show_package("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
-# package
-# 
-# # get all resources for this package
-# resources <- list_package_resources("c7d34d9b-23d2-44fe-8b3b-cd82c8b38978")
-# 
-# # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-# datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-# 
-# # load the first datastore resource as a sample
-# the_raw_data <- filter(datastore_resources, row_number()==1) %>% get_resource()
-# the_raw_data
-# 
-# #### Save data ####
-# write_csv(the_raw_data, "data/raw_data/raw_data.csv") 
 
 #Downloading all data from opendatatoronto directly to get all observations.
 all_data <- read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/c7d34d9b-23d2-44fe-8b3b-cd82c8b38978/resource/55136dac-26b1-4028-b9f5-7c2344f94153/download/bicycle-thefts%20-%204326.csv")

### 2024-09-27 scripts/02-data_cleaning.R
Edited preamble

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw data and add new variable to categorize bike cost.
 # Author: Taejun Um
-# Date: 24 September 2024
+# Date: 27 September 2024
 # Contact: taejun.um@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to have downloaded the data

### 2024-09-27 scripts/03-test_data.R
updated test script

@@ -1,19 +1,20 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
+# Purpose: Tests simulated and cleaned data of bicycle theft data
 # Author: Taejun Um
-# Date: 24 September 2024
+# Date: 27 September 2024
 # Contact: taejun.um@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to have simulated data
-# Any other information needed? None.s
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Test data ####
-data <- read_csv("data/raw_data/simulated.csv")
+# data <- read_csv("data/raw_data/simulated.csv")
+data <- read_csv("data/analysis_data/analysis_data.csv")
+
 
 #Test1 Check occ_date is in date form
 "Date" == class(data$occ_date)
@@ -21,13 +22,28 @@ data <- read_csv("data/raw_data/simulated.csv")
 #Test2 Check occ_year has a proper year
 all(format(data$occ_date, "%Y") == data$occ_year)
 
-#Test3 Check if premises_types have correct value
+#Test3 Check occ_month has a proper value
+all(data$occ_month %in% c("January", "February", "March", "April", "May", "June", 
+                          "July", "August", "September", "October", "November", "December"))
+
+#Test4 Check occ_hour has a proper value
+all(data$occ_hour %in% 1:23)
+
+#Test5 Check if premises_types have correct value
 premises_test <- c("Apartment", "Commercial", "Educational", "House", "Other",
                    "Outside", "Transit")
 all(data$premises_type %in% premises_test)
 
-#Test4 Check if bike_cost is correctly assigned for each observation
-data$bike_cost
+#Test6 Check if bike_cost is correctly assigned for each observation
+incorrect_rows <- data %>%
+  filter(
+    (bike_cost < 500 & cost_group != "< $500") |
+      (bike_cost >= 500 & bike_cost < 1000 & cost_group != "$500 < $1000") |
+      (bike_cost >= 1000 & bike_cost < 1500 & cost_group != "$1000 < $1500") |
+      (bike_cost >= 1500 & bike_cost < 2000 & cost_group != "$1500 < $2000") |
+      (bike_cost >= 2000 & cost_group != "$2000 <")
+  )
+nrow(incorrect_rows) == 0
 
-#Test5 Test for NAs
+#Test7 Test for NAs
 all(!is.na(data))

### 2024-09-27 scripts/00-simulate_data.R
Edited simulate script for typo

@@ -43,11 +43,11 @@ simulated_data <-
     bike_cost = round(rgamma(100, 3, 0.005), 2),
     cost_group =
       case_when(
-        bike_cost < 500 ~ "<500",
-        bike_cost < 1000 ~ "500<1000",
-        bike_cost < 1500 ~ "1000<1500",
-        bike_cost < 2000 ~ "1500<2000",
-        bike_cost >= 2000 ~ "2000<"
+        bike_cost < 500 ~ "< $500",
+        bike_cost < 1000 ~ "$500 < $1000",
+        bike_cost < 1500 ~ "$1000 < $1500",
+        bike_cost < 2000 ~ "$1500 < $2000",
+        bike_cost >= 2000 ~ "$2000 <"
       )
     )
 

### 2024-09-27 paper/paper.qmd
Editted format of paper

@@
 abstract: "In urban environments like Toronto, bicycle theft is as common as in any major city worldwide.
-This paper examines the trends and patterns of bicycle theft in Toronto from 2014 to 2022, focusing on the cost of bicycles and the types of locations where the thefts occurred.
-Data from this period show that the overall number of bicycle thefts did not significantly decline but began to decrease gradually from 2020, particularly in thefts occurring at residential properties and those involving high-end bicycles.
+This paper examines the trends and patterns of bicycle theft in Toronto from 2014 to 2023, focusing on the time of occurrence, the types of locations where the theft occurred, and the cost of bicycles.
+The following are the main findings from this data: the overall number of bicycle thefts has decreased since 2021, the an increasing trend of occurrences of incidents in both commercial districts and public transportation and there is a notable rise in high-end bicycles getting stolen.
 These findings can inform community efforts and assist Toronto's public authorities in developing more effective strategies to protect citizens' property and reduce bicycle theft."
@@
 
-This paper examines reported bicycle thefts in Toronto from 2014 to 2022.
+This paper examines reported bicycle thefts in Toronto from 2014 to 2023.
 The main objective is to observe and analyze the number of thefts, the locations of the incidents, and the price range of the stolen bicycles.
 Specifically, we will compare trends in the locations of incidents by year alongside changes in the price range of the targeted bicycles.
-Through this investigation, I confirmed that thefts steadily increased from 2014 until 2020, when the number of thefts began to decline.
+Through this investigation, I confirmed that thefts steadily increased from 2014 until 2020 but thefts dropped since 2021.
 Additionally, there have been notable changes in the locations where thefts occurred and the price range of the stolen bicycles over the years.
@@
 
-The remainder of this paper is structured as follows: @sec-data discusses the data and process of data cleaning, and @sec-first-point discusses our findings, limitation of the study, and suggestions to reduce bicycle theft.
+The remainder of this paper is structured as follows: @sec-data discusses the data and process of data cleaning, and @sec-discuss discusses our findings, limitations of the study, and suggestions to reduce bicycle theft.
 
@@
 
-The dataset used in the paper covers bicycle thefts reported to the Toronto Police system between 2014 and 2024.
-This dataset has been updated as of June 2024 and is considered reliable as it is based on the Toronto Police system.
+The dataset used in the paper covers reported bicycle thefts to the Toronto Police System between 2013 and 2024.
+This dataset was updated in June 2024 and is considered reliable as it is based on the Toronto Police system.
+Since the police system only has reported data, cases that are not reported to the police are not on the data.
+This data has variables of occurred time, reported time, stolen location, type of bike, cost of a bike, and current status of the bike.
 
 The dataset was downloaded from Toronto's Open Data Portal using opendatatoronto package [@opendatatoronto].
-Its setup was completed through R [@citeR] with the help of tidyverse package [@citetidyverse], janitor package [@citejanitor] and knitr package [@citeknitr].
+Its setup was completed through R [@citeR] with the help of tidyverse package [@citetidyverse], janitor package [@citejanitor], knitr package [@citeknitr], here package [@citehere], ggplot2 package [@citeggplot2], dplyr package [@citedplyr], tibble package [@citetibble], and readr package [@citereadr].
 
 The original dataset contained 28 variables, but I used the following: ID, occurrence date, occurrence year, occurrence month, occurrence hour, premises type, and bike cost.
-Then, I created a new group called "cost group" based on the bike cost, categorizing the bicycles according to their price range.
-Since the dataset is based on the date the theft was reported, I organized the data by the actual date of occurrence to reflect events from 2014 to 2022, excluding incidents that occurred before 2014.
+Then, I created a new group called "cost group" based on the bike cost, categorizing the bicycles with an increment of $500 according to their price range.
+Since the dataset is based on the date the theft was reported, I organized the data to reflect events from 2014 to 2023, excluding incidents that occurred before 2014 and after 2023.
 Additionally, during the data cleaning phase, observations with missing information were excluded.
-As a result, the dataset used in this paper contains 29,217 observations.
-(@tbl-head) shows an first 5 row of the data used.
+As a result, the dataset used in this paper contains 29,217 observations from 34,962 observations.
+(@tbl-head) shows the first 5 rows of the data used.
 
@@
 ## Number of Occurrences by Time
+### By Year
 
 First, we will examine how many bicycle thefts occurred by year.
-In @fig-occur-year-1, the line chart records thefts yearly and the table records the number of incidents in numbers.
+In @fig-occur-year, the line chart records thefts yearly and the table records the number of incidents in numbers.
 Bicycle thefts showed an increasing trend from 2014, peaking in 2018 with 3,418 recorded incidents, followed by a noticeable decrease starting in 2021.
 The decrements that happened in 2021 continued through 2023.
-According to @fig-occur-year-2, the number of incidents recorded in 2022 was 2,491 which was fewer than 2,511 incidents recorded in 2014.
+According to @tbl-occur-year, the number of incidents recorded in 2022 was 2,491 which was fewer than 2,511 incidents recorded in 2014.
 This indicates that the trend of bicycle theft has been decreasing over the last three years.
@@
 
+### By Month
+
 @fig-month shows the number of incidents that occurred by month with each line representing a different year.
@@
 
-### By hour
+### By Hour
 
@@
 @fig-hour-1 shows the average number of incidents across all years, while the @fig-hour-2 illustrates the number of incidents for each year over an hour.
-Graph 1 shows that there is a steady increase in thefts from 5 AM to 6 PM, and after that, the numbers begin to decline.
+@fig-hour-1 shows that there is a steady increase in thefts from 5 AM to 6 PM, and after that, the numbers begin to decline.
 Notably, there are three distinct peaks at 9 AM, 12 PM, and 6 PM, corresponding to periods of high social activity and movement.
@@
 In this dataset, crime scenes have been categorized into seven types: Apartment, Commercial, Educational, House, Outside, Transit, and Others.
-As seen in @fig-crime-scenes, there has been significant fluctuation in crimes committed in residential areas.
-Residential areas are classified into two categories: apartments and houses.
-Among them, apartment-related thefts showed a steady linear increase until 2019, but in 2020, the number dropped by around 200 incidents in one year, returning to 2017 levels by 2022.
-House-related thefts followed a zigzag pattern, increasing from 2014 and reaching a peak in 2020, before dropping to less than half by 2022.
-
-Another interesting point is that thefts committed outdoors were more frequent than those in apartments before 2019.
-However, starting in 2019, apartment thefts surpassed outdoor thefts, until 2022, when outdoor thefts once again became more common than apartment thefts.
+As seen in @fig-crime-scenes, the number of thefts shows a different pattern depending on the crime location.
+In the case of apartments, incidents surged from 315 in 2014 to 1,101 in 2020, marking a 249.5% increase over seven years.
+However, residential areas like apartments and houses peaked in 2020 and then declined.
+Apartment thefts dropped by 45.1% in just three years, with 604 incidents recorded in 2023, while house thefts saw a 42.3% decrease.
+'Outside' sector also experienced a peak in 2018, followed by a downward trend until 2023, with a 43.7% decline.
+'Commercial' buildings have shown a sinusoidal pattern since 2014. Although they hit a low point in 2021, thefts increased by 25.7% by 2023, indicating a recent and consistent upward trend.
+The 'Transit' sector, while having relatively fewer incidents compared to other sectors, has seen a rising trend since 2020.
+Starting with 40 incidents in 2020, the number nearly doubled to 78 by 2023, marking a 95% increase over three years.
+The 'Academic' and 'Other' sectors showed little variation in incident numbers by year. However, both sectors experienced declines in 2020 and 2022 respectively, followed by gradual increases.
 
@@
 
+premises_data_occ <- data %>%
+  group_by(occ_year, premises_type) %>%
+  summarise(number_of_occurrences = n(), .groups = "drop")
+
 ```
@@
 The original dataset includes a variable for the price of stolen bicycles, but it does not contain a variable for grouped prices.
-Therefore, I created a new variable called cost_group by grouping the prices in increments of $500.
+Therefore, I created a new variable called cost_group by grouping the prices in increments of $500 and it's divided into 5 groups.
 @fig-stolen-bike shows a bar graph with number of bike stolen by year categorized by cost group.
-From 2020 to 2022, the thefts of relatively inexpensive bicycles, priced under $1,000, have significantly decreased.
-On the other hand, high-end bicycles priced over $2,000 were stolen about 250 times in 2014.
-But this number has steadily increased, reaching approximately 500 thefts by 2022.
+The patterns of occurrence depend on the price range of the stolen bicycles.
+Bicycles under $1,000 have been on a downward trend.
+Bicycles under $500 and between $500 and $1,000 decreased 52.2 percent and 42.2 percent, respectively, in 2023 compared to 2017 and 2018.
+The number of stolen bicycles under $1,000 fell by about half from their peak, while high-end bicycles saw an increase in the number of occurrences.
+Bicycles priced between $1,000 and $1,500 saw a 64.4% increase in theft compared to 2014, while those priced above $2,000 saw a 119.4% increase compared to 2014.
+Bicycles ranging from $1500 to $2000 have recently seen a 34.8% increase from 2021 to 2023, although the number of thefts has not increased as much as other $1,000+ bikes.
+
 ```{r}
@@
   )
+cost_data_occ <- data %>%
+  group_by(occ_year, cost_group) %>%
+  summarise(number_of_occurrences = n(), .groups = "drop")
 ```
 
-# Discussion
+# Discussion {#sec-discuss}
 
-## Overall {#sec-first-point}
-The analysis of bicycle theft in Toronto has revealed trends and patterns over the nine years of theft incidents.
-All trends shifted starting in 2020, which appears to be a result of the lockdowns due to the pandemic.
-First, the overall number of thefts began to decrease from 2020, indicating a significant reduction in outdoor activities.
+## Overall
+The analysis of bicycle theft in Toronto has revealed trends and patterns over the ten years of theft incidents.
+Most of the trends shifted starting in 2021, which appears to be a result of the lockdowns due to the pandemic.
+First, the overall number of thefts began to decrease from 2021, indicating a significant reduction in outdoor activities.
 Bicycle owners were riding their bikes less frequently, and thieves were also unable to engage in outdoor activities due to the lockdowns.
 
-One of the most notable trends is that the daily theft graph exhibits a cyclical pattern.
+One of the most notable trends is that the monthly theft graph exhibits a cyclical pattern.
 This indicates that most crimes occur in the summer, while very few happen in the winter.
@@
 
-\appendix
-
-# Appendix {-}
-
-\newpage
-
 # References

### 2024-09-27 paper/paper.qmd
Edited minor errors in paper

@@
 As a result, the dataset used in this paper contains 29,217 observations from 34,962 observations.
-(@tbl-head) shows the first 5 rows of the data used.
+@tbl-head shows the first 5 rows of the data used.
 
@@
 #| label: fig-hour
-#| fig-cap: Number of bicycle theft incidents by hour in 10 years
-#| fig-subcap: ["Average in Year","Separated in Each Year"]
+#| fig-cap: Number of bicycle theft incidents by hour
+#| fig-subcap: ["Average by Year","Separated by Each Year"]
 #| warning: false
@@
   scale_color_brewer(palette = "Set1") +
-  labs(x = "Month", y = "Number of Occurrences", color = "Year") +
+  labs(x = "Hour", y = "Number of Occurrences", color = "Year") +
   theme_minimal() +
@@
 Therefore, I created a new variable called cost_group by grouping the prices in increments of $500 and it's divided into 5 groups.
-@fig-stolen-bike shows a bar graph with number of bike stolen by year categorized by cost group.
+@fig-group-cost shows a bar graph with number of bike stolen by year categorized by cost group.
 The patterns of occurrence depend on the price range of the stolen bicycles.
@@
 ```{r}
-#| label: fig-stolen-bike
-#| fig-cap: Total cost of bicycle by year
+#| label: fig-group-cost
+#| fig-cap: Number of occurrences categorized by price of bicycle
 #| fig-pos: "H"
@@
   facet_wrap(~ cost_group) +
-  labs(x = "Year", y = "Cost of Bicycle", fill = "Cost Group") +
+  labs(x = "Year", y = "Number of Occurrences", fill = "Cost Group") +
   theme(
